Server-Side Probabilistic Programming

### 2018-11-06 abstract/paper.tex
formatted

@@

\section{Challenges of Server-Side Probabilistic Programming}

@@

\paragraph{Simulation-inference dichotomy} Probabilistic models
often follow a design pattern of simulation-inference: a
significant part of the model is a simulator, running an
algorithm with fixed parameters; the optimal parameters, or
their distribution, are to be inferred. The inferred parameters
are then used by the software system to execute the simulation
independently of inference for forecasting and decision making.

@@

Table~\ref{tab:memory-runtime} provides memory consumption and
running time measurements on basic models to
illustrate  \texttt{infergo}'s performance.
The measurements were obtained on a 2.3GHz Intel Core 5
CPU with 8GB of memory for 1000 iterations of Hamiltonian Monte
Carlo with 10 leapfrog steps. Note that computation of log-likelihood
for standard distributions is not optimized yet. Quite the
opposite: since models in \texttt{infergo} are fully composable,
primitive distributions are themselves implemented as
\texttt{infergo} models and automatically differentiated.

\begin{table}[H]
	\caption{Memory and running times for 1000 iterations
	of HMC with 10 leapfrog steps.}
	\label{tab:memory-runtime}
	\begin{tabular}{r | c |  c | c}
		{\it model}  & \multicolumn{2}{c|}{\it time} & {\it memory} \\ 
		 & {\it compilation} & {\it execution} & \\\hline
		8 schools & 0.15s & 0.6s & 5.5MB \\
		10D normal, 100 points & 0.15s & 3.0s & 5.7MB \\
		50D normal, 100 points & 0.15s & 13.0s & 5.8MB 
	\end{tabular}
\end{table}

A lightweight probabilistic programming facility similar to
\texttt{infergo} can be added to most modern general-purpose
programming languages, in particular those used in implementing
large-scale software systems, making probabilistic
programming inference more accessible in server-side
applications.

>>>>

\linespread{1.03}

@@

\begin{sloppypar}
\section{Challenges of Server-Side Probabilistic Programming}

@@

\paragraph{Simulation vs. inference} Probabilistic models
often follow a design pattern of simulation-inference: a
significant part of the model is a simulator, running an
algorithm with fixed parameters; the optimal parameters, or
their distribution, are to be inferred. The inferred parameters
are then used by the software system to execute the simulation
independently of inference for forecasting and decision making.

@@

Table~\ref{tab:memory-runtime} provides memory consumption and
running time measurements on basic models to
illustrate  \texttt{infergo}'s performance.
The measurements were obtained on a 2.3GHz Intel Core 5
CPU with 8GB of memory for 1000 iterations of Hamiltonian Monte
Carlo with 10 leapfrog steps. Note that log-likelihood computation 
for standard distributions is not optimized yet. Quite the
opposite: since models in \texttt{infergo} are fully composable,
primitive distributions are themselves implemented as
\texttt{infergo} models and automatically differentiated.

{\smaller
\begin{table}[H]
	\caption{Memory and running times for 1000 iterations
	of HMC with 10 leapfrog steps.}
	\label{tab:memory-runtime}
	\begin{tabular}{r | c |  c | c}
		{\it model}  & \multicolumn{2}{c|}{\it time} & {\it memory} \\ 
		 & {\it compilation} & {\it execution} & \\\hline
		8 schools & 0.15s & 0.6s & 5.5MB \\
		10D normal, 100 points & 0.15s & 3.0s & 5.7MB \\
		50D normal, 100 points & 0.15s & 13.0s & 5.8MB 
	\end{tabular}
\end{table}}

A lightweight probabilistic programming facility similar to
\texttt{infergo} can be added to most modern general-purpose
programming languages, in particular those used in implementing
large-scale software systems, making probabilistic
programming inference more accessible in server-side
applications.
\end{sloppypar}
